---
title: "Onecut3 in mouse hypothalamus development"
author: "Evgenii Tretiakov"
date: "2021-01-28"
output: workflowr::wflow_html
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "#",
                      collapse = TRUE,
                      dev = c("png", "pdf"))
```
```{r load-packages}
# Load tidyverse infrastructure packages
library(here)
library(tidyverse)
library(magrittr)
library(zeallot)
library(future)

# Load packages for scRNA-seq analysis and visualisation
library(sctransform)
library(Seurat)
library(SeuratWrappers)
library(SeuratDisk)
library(UpSetR)
library(patchwork)
library(Nebulosa)
```
```{r set-paths}
src_dir    <- here("code")
data_dir   <- here("data")
output_dir <- here("output")
plots_dir  <- here(output_dir, "figures")
tables_dir <- here(output_dir, "tables")
```
```{r load-functions}
source(here(src_dir, "functions.R"))
source(here(src_dir, "genes.R"))
```
```{r set-seed}
reseed <- 42
set.seed(seed = reseed)
```
```{r set-parallel-execution-plan}
# available cores
n_cores <- available_cores(prop2use = .5)
# Parameters for parallel execution
plan("multicore", workers = n_cores)
options(future.globals.maxSize = Inf,
        future.rng.onMisuse = "ignore")
plan()
```

Read data
```{r load-data}
rar2020_ages_all     <- c("E15", "E17", "P00", "P02", "P10", "P23")
rar2020_ages_postnat <-                      c("P02", "P10", "P23")
samples_df <- read_tsv(here("data/samples.tsv"))
colours_wtree <- setNames(read_lines(here(data_dir, "colours_wtree.tsv")),
                          1:45)

onecut  <- LoadH5Seurat(here(data_dir, "rar2020.srt.cont.oc2or3.raw.h5seurat"))
onecut3 <- subset(onecut, subset = Onecut3 > 0)
```

Derive and filter matrix of Onecut3
```{r derive-mtxs}
mtx_oc3 <-
    onecut3 %>%
    GetAssayData("data", "RNA") %>%
    as.data.frame() %>%
    t()
rownames(mtx_oc3) <- colnames(onecut3)

# Filter features
filt_low_genes <-
    colSums(mtx_oc3) %>%
    .[. > quantile(., 0.4)] %>%
    names()
mtx_oc3 %<>% .[, filt_low_genes]

min_filt_vector <-
    mtx_oc3 %>%
    as_tibble() %>%
    select(all_of(filt_low_genes)) %>%
    summarise(across(.fns = ~ quantile(.x, .1))) %>%
    as.list %>%
    map(as.double) %>%
    simplify %>%
    .[colnames(mtx_oc3)]

# Prepare table of intersection sets analysis
content_mtx_oc3 <-
    (mtx_oc3 > min_filt_vector) %>%
    as_tibble() %>%
    mutate_all(as.numeric)
```

Correlation analysis visualisation between different genes
```{r plot-correlation-scatter}
p_corrs <- list(
    ggstatsplot::ggscatterstats(as.data.frame(mtx_oc3),
        x = Onecut3,  y = Trh, xfill = "#ffc400", yfill = "#e22ee2"),
    ggstatsplot::ggscatterstats(as.data.frame(mtx_oc3),
        x = Slc32a1,  y = Onecut3,  xfill = "#0000da",   yfill = "#ffc400"),
    ggstatsplot::ggscatterstats(as.data.frame(mtx_oc3),
        x = Th,  y = Onecut3,  xfill = "#006eff",   yfill = "#ffc400"),
    ggstatsplot::ggscatterstats(as.data.frame(mtx_oc3),
        x = Slc32a1,  y = Trh, xfill = "#0000da",   yfill = "#e22ee2"),
    ggstatsplot::ggscatterstats(as.data.frame(mtx_oc3),
        x = Th,  y = Trh, xfill = "#006eff",   yfill = "#e22ee2"),
    ggstatsplot::ggscatterstats(as.data.frame(mtx_oc3),
        x = Slc32a1,  y = Th,  xfill = "#0000da",   yfill = "#006eff"),
    ggstatsplot::ggscatterstats(as.data.frame(mtx_oc3),
        y = Slc32a1,  x = Onecut3,  yfill = "#0000da",   xfill = "#ffc400"),
    ggstatsplot::ggscatterstats(as.data.frame(mtx_oc3),
        y = Th,  x = Onecut3,  yfill = "#006eff",   xfill = "#ffc400"),
    ggstatsplot::ggscatterstats(as.data.frame(mtx_oc3),
        y = Slc17a6,  x = Onecut3,  yfill = "#ff0000",   xfill = "#ffc400"),
    ggstatsplot::ggscatterstats(as.data.frame(mtx_oc3),
        y = Gad1,  x = Onecut3,  yfill = "#a50202",    xfill = "#ffc400"),
    ggstatsplot::ggscatterstats(as.data.frame(mtx_oc3),
        y = Gad2,  x = Onecut3,  yfill = "#4002a5",    xfill = "#ffc400"),
    ggstatsplot::ggscatterstats(as.data.frame(mtx_oc3),
        y = Onecut2,  x = Onecut3,  yfill = "#6402a5",    xfill = "#ffc400"),
    ggstatsplot::ggscatterstats(as.data.frame(mtx_oc3),
        y = Nav1,  x = Onecut3,  yfill = "#2502a5",    xfill = "#ffc400"),
    ggstatsplot::ggscatterstats(as.data.frame(mtx_oc3),
        y = Nav2,  x = Onecut3,  yfill = "#4002a5",    xfill = "#ffc400"),
    ggstatsplot::ggscatterstats(as.data.frame(mtx_oc3),
        y = Trio,  x = Onecut3,  yfill = "#2502a5",    xfill = "#ffc400")
)
n_corrs <- list(
    "oc3-rna-data-Onecut3-Trh",
    "oc3-rna-data-Slc32a1-Onecut3",
    "oc3-rna-data-Th-Onecut3",
    "oc3-rna-data-Slc32a1-Trh",
    "oc3-rna-data-Th-Trh",
    "oc3-rna-data-Slc32a1-Th",
    "oc3-rna-data-Onecut3-Slc32a1",
    "oc3-rna-data-Onecut3-Th",
    "oc3-rna-data-Onecut3-Slc17a6",
    "oc3-rna-data-Onecut3-Gad1",
    "oc3-rna-data-Onecut3-Gad2",
    "oc3-rna-data-Onecut3-Onecut2",
    "oc3-rna-data-Onecut3-Nav1",
    "oc3-rna-data-Onecut3-Nav2",
    "oc3-rna-data-Onecut3-Trio"
)

walk2(n_corrs, p_corrs, save_my_plot, type = "stat-corr-plt")
```

Visualise intersections sets that we are going to use (highlighted)
```{r plot-upset-oc3}
upset(as.data.frame(content_mtx_oc3),
  order.by = "freq",
  sets.x.label = "Number of cells",
  text.scale = c(2, 1.6, 2, 1.3, 2, 3),
  nsets = 15,
  sets = c("Gad1", "Gad2", "Slc32a1", "Slc17a6"),
  queries = list(
    list(
      query = intersects,
      params = list("Gad1", "Gad2", "Slc32a1"),
      active = T
    ),
    list(
      query = intersects,
      params = list("Slc17a6"),
      active = T
    )
  ),
  nintersects = 60,
  empty.intersections = "on"
)
```

Regroup factor by stages for more balanced groups
```{r derive-stages}
onecut3$age %>% forcats::fct_count()
onecut3$stage <-
  onecut3$age %>%
  forcats::fct_collapse(`Embrionic day 15` = "E15",
                        `Embrionic day 17` = "E17",
                        Neonatal = c("P00", "P02"),
                        Postnatal = c("P10", "P23"))
onecut3$stage %>% forcats::fct_count()
```

Make subset of stable neurons
```{r derive-subset-def-cells}
onecut3$gaba_status <-
  content_mtx_oc3 %>%
  select(Gad1, Gad2, Slc32a1) %>%
  mutate(gaba = if_all(.fns = ~ .x > 0)) %>%
  .$gaba

onecut3$gaba_occurs <-
  content_mtx_oc3 %>%
  select(Gad1, Gad2, Slc32a1) %>%
  mutate(gaba = if_any(.fns = ~ .x > 0)) %>%
  .$gaba

onecut3$glut_status <-
  content_mtx_oc3 %>%
  select(Slc17a6) %>%
  mutate(glut = Slc17a6 > 0) %>%
  .$glut

oc3_fin <-
  subset(onecut3,
    cells = union(
      WhichCells(onecut3,
        expression = gaba_status == TRUE & glut_status == FALSE),
      WhichCells(onecut3,
        expression = glut_status == TRUE & gaba_occurs == FALSE)))
```

## Check contingency tables for neurotransmitter signature
```{r check-contingency-neurotransmitter-test}
oc3_fin@meta.data %>%
  janitor::tabyl(glut_status, gaba_status)
```

By age
```{r check-contingency-neurotransmitter-age}
oc3_fin@meta.data %>%
  janitor::tabyl(age, gaba_status)
```

By stage
```{r check-contingency-neurotransmitter-stage}
oc3_fin@meta.data %>%
  janitor::tabyl(stage, gaba_status)
```

Make splits of neurons by neurotransmitter signature 
```{r derive-subsets-by-neurotransmitter}
oc3_fin$status <- oc3_fin$gaba_status %>%
  if_else(true = "GABAergic",
    false = "glutamatergic")
Idents(oc3_fin) <- "status"
SaveH5Seurat(
  object    = oc3_fin,
  filename  = here(data_dir, "oc3_fin"),
  overwrite = TRUE,
  verbose   = TRUE
)

## Split on basis of neurotrans and test for difference
oc3_fin_neurotrans <- SplitObject(oc3_fin, split.by = "status")

## Split on basis of age and test for difference
oc3_fin_ages       <- SplitObject(oc3_fin, split.by = "age")
```

## DotPlots grouped by age
Expression of GABA receptors in GABAergic Onecut3 positive cells
```{r plot-dotplot-age-gabar-gaba}
DotPlot(object = oc3_fin_neurotrans$GABAergic,
        features = gabar,
        group.by = "age",
        cols = c("#adffff", "#0084ff"),
        col.min = -1, col.max = 1
) + RotatedAxis()
```

Expression of GABA receptors in glutamatergic Onecut3 positive cells
```{r plot-dotplot-age-gabar-glu}
DotPlot(object = oc3_fin_neurotrans$glutamatergic,
        features = gabar,
        group.by = "age",
        cols = c("#ffc2c2", "#ff3c00"),
        col.min = -1, col.max = 1
) + RotatedAxis()
```

Expression of glutamate receptors in GABAergic Onecut3 positive cells
```{r plot-dotplot-age-glur-gaba}
DotPlot(object = oc3_fin_neurotrans$GABAergic,
        features = glutr,
        group.by = "age",
        cols = c("#adffff", "#0084ff"),
        col.min = -1, col.max = 1
) + RotatedAxis()
```

Expression of glutamate receptors in glutamatergic Onecut3 positive cells
```{r plot-dotplot-age-glur-glu}
DotPlot(object = oc3_fin_neurotrans$glutamatergic,
        features = glutr,
        group.by = "age",
        cols = c("#ffc2c2", "#ff3c00"),
        col.min = -1, col.max = 1
) + RotatedAxis()
```

## DotPlots grouped by stage
Expression of GABA receptors in GABAergic Onecut3 positive cells
```{r plot-dotplot-stage-gabar-gaba}
DotPlot(object = oc3_fin_neurotrans$GABAergic,
        features = gabar,
        group.by = "stage",
        cols = c("#adffff", "#0084ff"),
        col.min = -1, col.max = 1
) + RotatedAxis()
```

Expression of GABA receptors in glutamatergic Onecut3 positive cells
```{r plot-dotplot-stage-gabar-glu}
DotPlot(object = oc3_fin_neurotrans$glutamatergic,
        features = gabar,
        group.by = "stage",
        cols = c("#ffc2c2", "#ff3c00"),
        col.min = -1, col.max = 1
) + RotatedAxis()
```

Expression of glutamate receptors in GABAergic Onecut3 positive cells
```{r plot-dotplot-stage-glur-gaba}
DotPlot(object = oc3_fin_neurotrans$GABAergic,
        features = glutr,
        group.by = "stage",
        cols = c("#adffff", "#0084ff"),
        col.min = -1, col.max = 1
) + RotatedAxis()
```

Expression of glutamate receptors in glutamatergic Onecut3 positive cells
```{r plot-dotplot-stage-glur-glu}
DotPlot(object = oc3_fin_neurotrans$glutamatergic,
        features = glutr,
        group.by = "stage",
        cols = c("#ffc2c2", "#ff3c00"),
        col.min = -1, col.max = 1
) + RotatedAxis()
```


```{r holder}

```