---
title: "Overrepresentation analysis of Th or Trh modulation in Onecut3 +/- subsets of glutamatergic neurons in the Lateral hypothalamic area dataset from Mickelsen LE et al 2019"
author: "Evgenii O. Tretiakov"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    df-print: paged
    code-fold: true
---

```{r setup, echo=FALSE, include=FALSE}
docname <- "onecut3_trh_contingency"
now <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
  if (before) {
    print(paste("Start:", Sys.time()))
    now <<- Sys.time()
  } else {
    print(paste("Stop:", Sys.time()))
    print(Sys.time() - now)
  }
})

knitr::opts_chunk$set(
  autodep        = TRUE,
  cache          = FALSE,
  cache.path     = paste0("cache/", docname, "/"),
  cache.comments = FALSE,
  cache.lazy     = FALSE,
  dev            = c("png", "pdf"),
  echo           = TRUE,
  error          = FALSE,
  fig.align      = "center",
  fig.width      = 14,
  fig.height     = 12,
  message        = FALSE,
  warning        = FALSE,
  timeit         = TRUE
)

# Load tidyverse infrastructure packages
suppressPackageStartupMessages({
  library(here)
  library(knitr)
  library(RColorBrewer)
  library(viridis)
  library(tidyverse)
  library(magrittr)
  library(stringr)
  library(skimr)
  library(future)
  library(zeallot)
  library(kableExtra)
  library(reticulate)
})

suppressPackageStartupMessages({
  library(Seurat)
  library(SeuratDisk)
  library(patchwork)
  library(UpSetR)
  library(scCustomize)
  library(Nebulosa)
})

# Set paths
src_dir <- here("code")
data_dir <- here("data")
output_dir <- here("output")
plots_dir <- here(output_dir, "figures")
tables_dir <- here(output_dir, "tables")
source(here(src_dir, "genes.R"))
source(here(src_dir, "functions.R"))

# parallelisation
n_cores <- 32

# set seed
reseed <- 42
set.seed(seed = reseed)

# ggplot2 theme
theme_set(ggmin::theme_powerpoint())
```


```{r params}
bioproject <- "PRJNA515063"
project <- "mickelsen2019_LHA"
cb_fpr <- 0.001
low_cutoff_gene <- 500
high_cutoff_gene <- NULL
high_cutoff_gene <- 7000
low_cutoff_umis <- NULL
low_cutoff_umis <- -Inf
high_cutoff_umis <- 25000
high_cutoff_pc_mt <- 15
high_cutoff_pc_ribo <- 15
high_cutoff_pc_hb <- 0.1
high_cutoff_doublet_score <- 0.33
high_cutoff_complexity <- 0.85
```


```{r load-data}
srt <- LoadH5Seurat(
  file = here(
    data_dir,
    sprintf("%s-whole_dataset-fpr_%s-clusters.h5Seurat", bioproject, cb_fpr)
  )
)
srt
neurons <- subset(srt, subset = Rbfox3 > 0 | Elavl4 > 0 | Snap25 > 0 | Stmn2 > 0)
onecut <- subset(neurons, subset = Onecut3 > 0 | Onecut2 > 0)
onecut3 <- subset(onecut, subset = Onecut3 > 0 & Slc17a6 > 0)
```

```{r}
mtx_neurons <-
  onecut3 %>%
  GetAssayData("data", "RNA") %>%
  as.data.frame() %>%
  t()
rownames(mtx_neurons) <- colnames(onecut3)

# Filter features
filt_low_genes <-
  colSums(mtx_neurons) %>%
  .[. > quantile(., 0.2)] %>%
  names()
mtx_neurons %<>% .[, filt_low_genes]

min_filt_vector <-
  mtx_neurons %>%
  as_tibble() %>%
  select(all_of(filt_low_genes)) %>%
  summarise(across(.fns = ~ quantile(.x, .1))) %>%
  as.list() %>%
  map(as.double) %>%
  simplify() %>%
  .[colnames(mtx_neurons)]

# Prepare table of intersection sets analysis
content_mtx_neuro <-
  (mtx_neurons > min_filt_vector) %>%
  as_tibble() %>%
  mutate_all(as.numeric)

onecut3$trh_status <-
  content_mtx_neuro %>%
  select(Trh) %>%
  mutate(trh = Trh > 0) %>%
  .$trh
onecut3$trh_status <- onecut3$trh_status %>%
  if_else(true = "Trh",
    false = "query"
  )

## Split on basis of neurotrans and test for difference
onecut3_trh <- SplitObject(onecut3, split.by = "trh_status")

markers_logreg <-
  FindAllMarkers(
    onecut3,
    assay = "RNA",
    verbose = FALSE,
    random.seed = reseed,
    only.pos = TRUE,
    min.pct = 0.1,
    base = 10,
    logfc.threshold = 0.1,
    densify = TRUE,
    test.use = "LR"
  )
write_csv(
  markers_logreg,
  here(
    tables_dir, "logreg-Vglut2_Onecut3_populations-DEG.csv"
  )
)
```

```{r}
top10_markers <-
  Extract_Top_Markers(
    marker_dataframe = markers_logreg,
    num_genes = 10,
    named_vector = FALSE,
    make_unique = TRUE,
    rank_by = "avg_log10FC"
  )

Iterate_FeaturePlot_scCustom(
  seurat_object = onecut3,
  gene_list = top10_markers,
  single_pdf = TRUE,
  reduction = "umap",
  colors_use = viridis(
    n = 30,
    alpha = .55,
    direction = -1,
    option = "E"
  ),
  pt.size = 3,
  alpha_exp = 0.45,
  alpha_na_exp = 0.1,
  file_path = here(plots_dir),
  file_name = "logreg_top10-Vglut2_Onecut3_populations-DEG.pdf"
)
```


## DotPlots grouped by orig.ident
Expression of GABA receptors in query Onecut3 positive cells
```{r plot-dotplot-orig.ident-gabar-query, fig.height=3, fig.width=9}
DotPlot(
  object = onecut3_trh$query,
  features = gabar,
  group.by = "orig.ident",
  cols = c("#adffff", "#0084ff"),
  assay = "RNA"
) + RotatedAxis()
```

Expression of GABA receptors in Trh Onecut3 positive cells
```{r plot-dotplot-orig.ident-gabar-trh, fig.height=3, fig.width=9}
DotPlot(
  object = onecut3_trh$Trh,
  features = gabar,
  group.by = "orig.ident",
  cols = c("#ffc2c2", "#ff3c00"),
  assay = "RNA"
) + RotatedAxis()
```

Expression of glutamate receptors in query Onecut3 positive cells
```{r plot-dotplot-orig.ident-glur-query, fig.height=3, fig.width=9}
DotPlot(
  object = onecut3_trh$query,
  features = glutr,
  group.by = "orig.ident",
  cols = c("#adffff", "#0084ff"),
  assay = "RNA"
) + RotatedAxis()
```

Expression of glutamate receptors in Trh Onecut3 positive cells
```{r plot-dotplot-orig.ident-glur-trh, fig.height=3, fig.width=9}
DotPlot(
  object = onecut3_trh$Trh,
  features = glutr,
  group.by = "orig.ident",
  cols = c("#ffc2c2", "#ff3c00"),
  assay = "RNA"
) + RotatedAxis()
```

## DotPlots
Expression of GABA receptors in query Onecut3 positive cells
```{r plot-dotplot-gabar-query, fig.height=5, fig.width=9}
DotPlot(
  object = onecut3_trh$query,
  features = gabar,
  cols = c("#adffff", "#0084ff"),
  assay = "RNA"
) + RotatedAxis()
```

Expression of GABA receptors in Trh Onecut3 positive cells
```{r plot-dotplot-gabar-trh, fig.height=3, fig.width=9}
DotPlot(
  object = onecut3_trh$Trh,
  features = gabar,
  cols = c("#ffc2c2", "#ff3c00"),
  assay = "RNA"
) + RotatedAxis()
```

Expression of glutamate receptors in query Onecut3 positive cells
```{r plot-dotplot-glur-query, fig.height=5, fig.width=9}
DotPlot(
  object = onecut3_trh$query,
  features = glutr,
  cols = c("#adffff", "#0084ff"),
  assay = "RNA"
) + RotatedAxis()
```

Expression of glutamate receptors in Trh Onecut3 positive cells
```{r plot-dotplot-glur-trh, fig.height=3, fig.width=9}
DotPlot(
  object = onecut3_trh$Trh,
  features = glutr,
  cols = c("#ffc2c2", "#ff3c00"),
  assay = "RNA"
) + RotatedAxis()
```

## Neuropeptides

Expression of Neuropeptides in query Onecut3 positive cells
```{r plot-dotplot-np-query, fig.height=5, fig.width=11.5}
DotPlot(
  object = onecut3_trh$query,
  features = np,
  cols = c("#adffff", "#0084ff"),
  assay = "RNA"
) + RotatedAxis()
```

Expression of Neuropeptides in Trh Onecut3 positive cells
```{r plot-dotplot-np-trh, fig.height=3, fig.width=11.5}
DotPlot(
  object = onecut3_trh$Trh,
  features = np,
  cols = c("#ffc2c2", "#ff3c00"),
  assay = "RNA"
) + RotatedAxis()
```

Expression of Neuropeptides receptors in query Onecut3 positive cells
```{r plot-dotplot-npr-query, fig.height=5, fig.width=11.5}
DotPlot(
  object = onecut3_trh$query,
  features = npr,
  cols = c("#adffff", "#0084ff"),
  assay = "RNA"
) + RotatedAxis()
```

Expression of Neuropeptides receptors in Trh Onecut3 positive cells
```{r plot-dotplot-npr-trh, fig.height=3, fig.width=11.5}
DotPlot(
  object = onecut3_trh$Trh,
  features = npr,
  cols = c("#ffc2c2", "#ff3c00"),
  assay = "RNA"
) + RotatedAxis()
```

```{r drv-sbs-mtx}
sbs_mtx_oc <-
  onecut %>%
  GetAssayData("data", "RNA") %>%
  as.data.frame() %>%
  t()
rownames(sbs_mtx_oc) <- colnames(onecut)

# Filter features
filt_low_genes2 <-
  colSums(sbs_mtx_oc) %>%
  .[. > quantile(., 0.4)] %>%
  names()
sbs_mtx_oc %<>% .[, filt_low_genes2]

min_filt_vector2 <-
  sbs_mtx_oc %>%
  as_tibble() %>%
  select(all_of(filt_low_genes2)) %>%
  summarise(across(.fns = ~ quantile(.x, .005))) %>%
  as.list() %>%
  map(as.double) %>%
  simplify() %>%
  .[filt_low_genes2]

# Prepare table of intersection sets analysis
content_sbs_mtx_oc <-
  (sbs_mtx_oc > min_filt_vector2) %>%
  as_tibble() %>%
  mutate_all(as.numeric)


onecut$gaba_status <-
  content_sbs_mtx_oc %>%
  select(Gad1, Gad2, Slc32a1) %>%
  mutate(gaba = if_all(.fns = ~ .x > 0)) %>%
  .$gaba

onecut$gaba_occurs <-
  content_sbs_mtx_oc %>%
  select(Gad1, Gad2, Slc32a1) %>%
  mutate(gaba = if_any(.fns = ~ .x > 0)) %>%
  .$gaba

onecut$glut_status <-
  content_sbs_mtx_oc %>%
  select(Slc17a6) %>%
  mutate(glut = Slc17a6 > 0) %>%
  .$glut

oc_fin <-
  subset(onecut,
    cells = union(
      WhichCells(onecut,
        expression = gaba_status == TRUE & glut_status == FALSE
      ),
      WhichCells(onecut,
        expression = glut_status == TRUE & gaba_occurs == FALSE
      )
    )
  )

oc_fin$status <- oc_fin$gaba_status %>%
  if_else(true = "query",
    false = "Trh"
  )
Idents(oc_fin) <- "status"

sbs_mtx_oc <-
  oc_fin %>%
  GetAssayData("data", "RNA") %>%
  as.data.frame() %>%
  t()
rownames(sbs_mtx_oc) <- colnames(oc_fin)

# Filter features
filt_low_genes2 <-
  colSums(sbs_mtx_oc) %>%
  .[. > quantile(., 0.4)] %>%
  names()
sbs_mtx_oc %<>% .[, filt_low_genes2]

min_filt_vector2 <-
  sbs_mtx_oc %>%
  as_tibble() %>%
  select(all_of(filt_low_genes2)) %>%
  summarise(across(.fns = ~ quantile(.x, .005))) %>%
  as.list() %>%
  map(as.double) %>%
  simplify() %>%
  .[filt_low_genes2]

# Prepare table of intersection sets analysis
content_sbs_mtx_oc <-
  (sbs_mtx_oc > min_filt_vector2) %>%
  as_tibble() %>%
  mutate_all(as.numeric)
```

```{r plot-upset-sbs-oc3}
upset(as.data.frame(content_sbs_mtx_oc),
  order.by = "freq",
  sets.x.label = "Number of cells",
  text.scale = c(2, 1.6, 2, 1.3, 2, 3),
  nsets = 15,
  sets = c("Gad1", "Gad2", "Slc32a1", "Slc17a6", "Onecut3"),
  queries = list(
    list(
      query = intersects,
      params = list("Gad1", "Gad2", "Slc32a1", "Onecut3"),
      active = T
    ),
    list(
      query = intersects,
      params = list("Slc17a6", "Onecut3"),
      active = T
    )
  ),
  nintersects = 20,
  empty.intersections = NULL
)
```

```{r plot-upset-sbs-oc3-mod}
upset(as.data.frame(content_sbs_mtx_oc),
  order.by = "freq",
  sets.x.label = "Number of cells",
  text.scale = c(2, 1.6, 2, 1.3, 2, 3),
  nsets = 15,
  sets = c("Th", "Trh", "Slc32a1", "Slc17a6", "Onecut3"),
  queries = list(
    list(
      query = intersects,
      params = list("Th", "Slc32a1", "Onecut3"),
      active = T
    ),
    list(
      query = intersects,
      params = list("Trh", "Slc17a6", "Onecut3"),
      active = T
    )
  ),
  nintersects = 20,
  empty.intersections = NULL
)
```

```{r plot-piecharts-sbs-oc3-chisqr, fig.height=10, fig.width=12}
sbs_mtx_oc_full <- content_sbs_mtx_oc |>
  select(any_of(c(neurotrans, glutr, gabar, "Trh", "Th", "Onecut3"))) |>
  dplyr::bind_cols(oc_fin@meta.data)

sbs_mtx_oc_full |> glimpse()

sbs_mtx_oc_full$modulator <-
  sbs_mtx_oc_full %>%
  select(Trh, Th) %>%
  mutate(modulator = if_any(.fns = ~ .x > 0)) %>%
  .$modulator

sbs_mtx_oc_full$oc3 <-
  (sbs_mtx_oc_full$Onecut3 > 0)

library(ggstatsplot)
# for reproducibility
set.seed(123)

## plot
# ggpiestats(
#   data = sbs_mtx_oc_full,
#   x = modulator,
#   y = status,
#   perc.k = 1,
#   package = "ggsci",
#   palette = "category10_d3",
# ) + # further modification with `{ggplot2}` commands
#   ggplot2::theme(
#     plot.title = ggplot2::element_text(
#       color = "black",
#       size = 14,
#       hjust = 0
#     )
#   )

# plot
grouped_ggpiestats(
  # arguments relevant for `ggpiestats()`
  data = sbs_mtx_oc_full,
  x = modulator,
  y = oc3,
  grouping.var = status,
  perc.k = 1,
  package = "ggsci",
  palette = "category10_d3",
  # arguments relevant for `combine_plots()`
  title.text = "Neuromodulator specification of onecut-driven hypothalamic neuronal lineages by Onecut3 and main neurotransmitter expression",
  caption.text = "Asterisks denote results from proportion tests; \n***: p < 0.001, ns: non-significant",
  plotgrid.args = list(nrow = 2)
)
```

## Session information

```{r session-info, cache = FALSE}
devtools::session_info()
```
